---
title: "Grade Analysis"
author: Daniel Rohrbach
format: html
editor: visual
echo: false
message: false
warning: false
---

```{r}
here::i_am("r-101-grade-analysis.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(vroom)
library(stringr)
```

https://github.com/dauphine-data-daniel/r-101-grade-analysis

## Question 1
Data set got loaded.

```{r}
grades <- vroom(here("grades.csv"))
```

## Question 2

```{r}
exam_stats <- grades |> 
  summarise(
    Min = min(Exam, na.rm = TRUE),
    Max = max(Exam, na.rm = TRUE),
    Median = median(Exam, na.rm = TRUE),
    Mean = mean(Exam, na.rm = TRUE)
  )

knitr::kable(exam_stats, caption = "Exam Statistics", align = 'c')
```

## Question 3

```{r}
missing_exam <- grades |> 
  filter(is.na(Exam)) |> 
  count()
```

We observe that `r missing_exam` students did not take the final exam.

## Question 4

```{r}
grades |> 
  filter(!is.na(Exam)) |> 
  ggplot(aes(x = Exam)) + 
  geom_histogram(binwidth = 1, fill = "lightblue", color = "white") +
  labs(title = "Distribution of Exam Grades", 
       x = "Grade", 
       y = "Number of Students") +
  theme_minimal()
```

For reasons of graphical representation, .5 grades were rounded to full grades using the "binwidth=1" command.

## Question 5

```{r}
group_counts <- grades |> 
  group_by(Group) |> 
  summarise(Count = n())

knitr::kable(group_counts, caption = "Number of Students per Group", align = 'c')
```

## Question 6

```{r}
grades |>
  group_by(Group) |>
  summarise(count = n()) |>
  ggplot(aes(x = Group, y = count)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "white") + 
  labs(title = "Number of Students in Each Group", 
       x = "Group", 
       y = "Number of Students") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## Question 7

```{r}
grades |>
  ggplot(aes(x = Group, y = Exam)) + 
  geom_boxplot(aes(fill = Group)) + 
  labs(title = "Distribution of Exam Grades by Group (Boxplot)", 
       x = "Group", 
       y = "Exam Grade") +
  theme_minimal() + 
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r}
grades |>
  ggplot(aes(x = Group, y = Exam)) + 
  geom_violin(aes(fill = Group)) + 
  labs(title = "Distribution of Exam Grades by Group (Violin Plot)", 
       x = "Group", 
       y = "Exam Grade") +
  theme_minimal() + 
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## Question 8

```{r}
missing_exams_per_group <- grades |> 
  group_by(Group) |> 
  summarise(missing_exams = sum(is.na(Exam))) |>
  filter(missing_exams > 0) |>
  ungroup()

knitr::kable(missing_exams_per_group, 
             caption = "Number of Students Missing Exam by Group", 
             col.names = c("Group", "Number Missing Exam"),
             align = c('l', 'c'))
```

## Question 9
```{r}
missing_exams_per_group <- grades |> 
  group_by(Group) |> 
  summarise(missing_exams = sum(is.na(Exam), na.rm = TRUE)) |>
  ungroup()

ggplot(missing_exams_per_group, aes(x = Group, y = missing_exams, fill = Group)) +
  geom_col() +
  labs(title = "Number of Students Missing Exam by Group",
       x = "group",
       y = "amount of students who missed the exam") +
  scale_fill_brewer(palette = "Paired") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Question 10
Creation of a new reshaped data frame. Id and Group are excluded from reshaping.
```{r}
grades_long <- grades |> 
  pivot_longer(
    cols = -c(Id, Group),
    names_to = "name",
    values_to = "value"
  )
```

## Question 11 
Using the long format, the number of missing grades in total for each student is computed.
```{r}
missing_grades_per_student <- grades_long |> 
  group_by(Id) |> 
  summarise(missing_grades = sum(is.na(value)))
```
## Question 12
```{r}
ggplot(missing_grades_per_student, aes(x = missing_grades)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "white") +
  labs(title = "Distribution of Missing Grades per Student",
       x = "Number of Missing Grades",
       y = "Count of Students") +
  theme_minimal()
```
## Question 13
```{r}
missing_exams_per_group <- grades_long |>
  filter(str_starts(name, "Exam")) |>
  group_by(Group) |>
  summarise(missing_exams = sum(is.na(value), na.rm = TRUE)) |>
  ungroup()

knitr::kable(missing_exams_per_group, 
             caption = "Number of Students Who Missed the Exam by Group")
```
## Question 14
The Number of Missing Online Tests per student is computed.
```{r}
missing_online_tests_per_student <- grades_long |>
  filter(str_starts(name, "Online_MCQ")) |>
  group_by(Id) |>
  summarise(missing_online_tests = sum(is.na(value), na.rm = TRUE)) |>
  ungroup()

```

## Question 15
```{r}
missing_online_tests <- grades |> 
  rowwise() |> 
  mutate(missing_online_tests = sum(is.na(c_across(starts_with("Online_MCQ"))), na.rm = TRUE)) |> 
  select(Id, missing_online_tests)

grades_with_missing_online_tests <- grades |> 
  left_join(missing_online_tests, by = "Id")

ggplot(grades_with_missing_online_tests, aes(x = missing_online_tests, y = Exam)) +
  geom_bin2d(bins = 19, aes(fill = ..count..)) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", guide = "colourbar") + # Colors for low and high counts
  labs(
    x = "Number of Missing Online Tests",
    y = "Exam Grade",
    title = "Exam Grades Conditioned by Number of Missing Online Tests"
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## Question 16
Table got created with two columns "Id" and "Missed" with value TRUE when the student misses at least one "MCQ_xx" grade and FALSE when they miss no grade.
```{r}
missed_mcqs <- grades |>
  rowwise() |>
  mutate(Missed = any(is.na(c_across(starts_with("MCQ_"))))) |>
  select(Id, Missed) |>
  ungroup()
```

## Question 17
Table got created with two columns "Group" and "P_missed" with percentage of students in each group who missed at least one "MCQ_xx grade".
```{r}
grades_with_missed_mcq <- grades |>
  rowwise() |>
  mutate(Missed = any(is.na(c_across(starts_with("MCQ_"))))) |>
  ungroup()

percentage_missed_per_group <- grades_with_missed_mcq |>
  group_by(Group) |>
  summarise(P_missed = mean(Missed) * 100) |>
  ungroup()
```

## Question 18
```{r}
average_exam_grade_per_group <- grades |>
  group_by(Group) |>
  summarise(Average_Exam_Grade = mean(Exam, na.rm = TRUE)) |>
  ungroup()

merged_data <- inner_join(average_exam_grade_per_group, percentage_missed_per_group, by = "Group")

ggplot(merged_data, aes(x = P_missed, y = Average_Exam_Grade)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Percentage of Missed MCQ Grades",
       y = "Average Exam Grade",
       title = "Average Exam Grade per Group vs. Percentage of Missed MCQ Grades") +
  theme_minimal()

```



